Learning and calibrating per-location classifiers for visual place recognition\thanks{

### 2015-11-24 manuscript.tex
w-norm calibration

\begin{abstract}
  The aim of this work is to localize a query photograph by finding other images depicting the same place in a large geotagged image database. This is a challenging task due to changes in viewpoint, imaging conditions and the large size of the image database. The contribution of this work is two-fold. First, we cast the place recognition problem as a classification task and use the available geotags to train a classifier for each location in the database in a similar manner to per-exemplar SVMs in object recognition. 
  Second, as only a few positive training examples are available for each location, we propose 
  two methods to calibrate all the per-location SVM classifiers without the need for additional positive training data.
  The first method relies on p-values from statistical hypothesis testing and uses only the available negative training data. The second method performs an affine calibration by appropriately normalizing the learnt classifier hyperplane and does not need any additional labelled training data. We test the proposed place recognition method with the bag-of-visual-words and Fisher vector image representations suitable for large scale indexing. 
  {  
    Experiments are performed on three datasets: 25,000 and 55,000 geotagged street view images of Pittsburgh, and the 24/7 Tokyo benchmark containing $76,000$ images with varying illumination conditions. The results show improved place recognition accuracy of the learnt image representation over direct matching of raw image descriptors.
  }

@@

  In this work, we also take advantage of geotags as an available form of supervision and investigate whether the place recognition problem can be cast as a classification task.
  {
    Learning visual classifiers has been investigated for landmark recognition~\cite{Li09} where consumer photographs were clustered into landmark classes based on geo-tags. 
    In this work we wish to recognize individual street locations rather than a small number of landmarks, and as a consequence have only a few (1-5) photographs available for each place.  In particular, we train a classifier {\em for each location on the map} in a similar manner to per-exemplar classification in object recognition~\cite{Malisiewicz11}. 
  }

@@

  While learning classifiers for each place may be appealing, calibrating outputs of the individual classifiers is a critical issue. In object recognition~\cite{Malisiewicz11}, it is addressed in a separate calibration stage on a held-out set of training data.
  This is not possible in the place recognition set-up as only a small number, typically one to five, of positive training images are available for each location (e.g. Street View images viewing the same building facade). To address this issue, we propose two calibration methods. 
  The first method relies on p-values from statistical hypothesis testing and uses only the available negative training data. The second method performs a simple affine calibration by appropriately normalizing the learnt classifiers and does not need any additional labelled calibration examples.

@@

  {
     The task of geo-localizing a given input query photograph has recently received considerable attention. The output can be a coarse geo-localization on the level of continents and cities~\cite{Doersch12,Hays2008,Kalogerakis09} or a name of the depicted landmark~\cite{Li09}.  In this work we focus on visually recognizing the ``same place" by finding an image in geo-tagged database that depicts the same building facade or street-corner as shown in the query~\cite{Chen11,Cummins09,Knopp2010,Schindler07,Torii2013,Zamir10}.
  }

  This {\em visual place recognition} problem is typically treated as large-scale instance-level retrieval~\cite{Cummins09,Chen11,Knopp2010,Schindler07,Torii2013,Zamir10}, where images are represented using local invariant features~\cite{Lowe04} encoded and aggregated into the bag-of-visual-words~\cite{Csurka04,Sivic03} or Fisher vector~\cite{Jegou12} representations. The image database can be further augmented by 3D point clouds~\cite{Klinger13}, automatically reconstructed by large-scale structure from motion (SfM)~\cite{Agarwal-ICCV-2009,Klinger13}, which enables accurate prediction of query image camera position~\cite{Li12,Sattler12}.
  In contrast, in this work we investigate learning a discriminative place-specific image representation. A similar idea has been recently explored in~\cite{Cao13} who learn a graph-based discriminative representation for landmark image collections where typically many images are available for each landmark.
  In this work, we focus on street-level images such as Google Street View, which have greater coverage, but typically only one or a small number of images are  available for each place.  To address this issue we learn a discriminative re-weighting of the descriptor specific to each image in the database using per-exemplar support vector machine~\cite{Malisiewicz11}.

  \paragraph{Per-exemplar support vector machines.} 
    The exemplar support vector machines (e-SVM) has been used in a number of visual recognition tasks including category-level recognition~\cite{Malisiewicz11}, cross-domain retrieval~\cite{Shrivastava11}, scene parsing~\cite{Tighe13} or as an initialization  for more complex discriminative clustering models~\cite{Doersch12,Singh12}. The main idea is to train a linear support vector machine (SVM) classifier from a single positive example and a large number of negatives. The intuition is that the resulting weight vector will give a higher weight to the discriminative dimensions of the positive training data point and will down weight dimensions that are non-discriminative with respect to the negative training data.

    {
      The exemplar support vector machine can be learnt at query time where the weight vector is used as a new query image representation~\cite{Shrivastava11}. However, this requires training a new classifier afresh for each query that is computationally very demanding. In this work, similar to~\cite{Malisiewicz11} who learn per-exemplar object category representation, we learn per-exemplar classifiers for each place in the database off-line.  A key advantage is that each per-exemplar classifier is trained independently and hence the learning can be heavily parallelized. The per-exemplar training brings, however, also an important drawback. As each classifier is trained independently a careful calibration of the resulting classifier scores is required~\cite{Malisiewicz11}.
    }

@@

\paragraph{Contributions.} 
  This paper has two main contributions. 
  {
    First, we cast the place recognition problem as a classification task where we use the available geo-tags as a weak form of supervision to train a classifier for each location in the database (section~\ref{sec:classifiers}).
    These classifiers are subsequently used for ranking the database images at query time.
  }

  Second, as only a few positive training examples are available for each location, we propose two methods to calibrate all the per-location SVM classifiers without the need for additional positive training data. The first method (section~\ref{sec:calibration}) relies on p-values from statistical hypothesis testing. The second method (section~\ref{sec:calibrationRenorm}) performs an affine calibration by appropriately normalizing the learnt decision hyperplane. We also describe a memory efficient classifier representation for the sparse bag-of-visual-word vectors (section~\ref{sec:memory}) and  experimentally demonstrate benefits of the proposed approach (section~\ref{sec:experiments} and~\ref{sec:results}).

%%%%%%%%%%%%%%%%%%%%%%%%
\section{Per-location classifiers for place recognition}
\label{sec:classifiers}
%%%%%%%%%%%%%%%%%%%%%%%%
   We are given an image descriptor $\vec{x}_j$, one for each database image $j$. This representation can be a sparse tf-idf weighted bag-of-visual-words vector~\cite{Sivic03} or a dense compact descriptor such as the Fisher vector (FV)~\cite{Jegou12}. The goal is to learn a score $f_j$ for each database image $j$, so that, at test time, given the descriptor $\vec{q}$ of the query image, we can either retrieve the correct target image as the image $j^*$ with the highest score
   % 
   \begin{equation}
   \label{eq:class}
    j^*=\operatorname*{arg\;max}_{j} f_j(\vec{q}) 
   \end{equation}
    %
   \noindent
   or use these scores to rank candidate images and use geometric verification to identify the correct location in an $n$-best list.
   Instead of approaching the problem directly as a large multiclass classification problem, we tackle the problem by learning a per-exemplar linear SVM classifier~\cite{Malisiewicz11}  for each database image $j$.
   Similar to~\cite{Knopp2010}, we use the available geotags to construct the negative set $\mathcal N_j$ for each image $j$. The negative set is constructed so as to concentrate difficult negative examples, i.e.\ from images that are far away from the location of image $j$ and similar to the target image as measured by the dot product between their feature vectors. The details of the construction procedure will be given in section~\ref{sec:experiments}.  The positive set $\mathcal P_j$ is represented o single positive example, which is $\vec{x}_j$ itself. 
   Each SVM classifier produces a score $s_j$ which is a priori not comparable with the score of the other classifiers. A calibration of these scores will therefore be key to convert them to comparable scores $f_j$. This calibration problem is more difficult than usual given that we only have a single positive example and will be addressed in section~\ref{sec:calibration}.

   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
   \subsection{Learning per-location SVM classifiers }
   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
      Each linear SVM classifier %learns 
      {generates} a score $s_j$ of the form 
      %
      \begin{align}
        s_j(\vec{q})=\vec{q}^T \vec{w}_j+b_j
        \label{eq:linear}
      \end{align}
      %
      \noindent
      where $\vec{w}_j$ is a weight vector re-weighting contributions of individual visual words and $b_j$ is the bias specific for image $j$. Given the training sets $\mathcal P_j$ and $\mathcal N_j$, the aim is to find a vector $\vec{w}_j$ and bias $b_j$ such that the score difference between $\vec{x_j}$ and the closest neighbor from its negative set $\mathcal N_j$ is {\em maximized}.
      Learning the weight vector $\vec{w}_j$ and bias $b_j$ is formulated as a minimization of the convex objective 
      %
      \begin{align}
        \nonumber
        \Omega(\vec{w}_j,b_j)=||\vec{w}_j||^{2}& +C_1\sum_{\vec{x}\in \mathcal P_j}h(\vec{w}_j^T\vec{x}+b_j)   \\
        \label{eq:obj}
                           & +C_2\sum_{\vec{x}\in \mathcal N_j}h(-\vec{w}_j^T\vec{x}-b_j), 
      \end{align}
      %
      \noindent
      where the first term is the regularizer, the second term is the loss on the positive training data weighted by scalar parameter $C_1$, and the third term is the loss on the negative training data weighted by scalar parameter $C_2$.   
      This is a standard SVM formulation \eqref{eq:obj}, also used in exemplar-SVM~\cite{Malisiewicz11}.
      In our case $h$ is the squared hinge loss, which we found to work better in our setting than the standard hinge-loss. Parameters $\vec{w}_j$ and $b_j$ are learned separately for each database image $j$ in turn.

@@

      Two other examples of score calibration functions are shown in figure~\ref{fig:3qVSw} in section~\ref{sec:results}. Note that while figure~\ref{fig:calib} illustrates only few points on the cdf, the two plots in figure~\ref{fig:3qVSw} show a complete cdf that contains on the order of $25k$ data points. Note also that the two cumulative density functions in figure~\ref{fig:3qVSw} are similar but not identical.

@@

  \subsection{Calibration by normalization}
  \label{sec:calibration_by_renormalization}
    Parameters of the affine calibration function~\eqref{eq:affine} could be learnt from negative training data in a similar manner to, for example,~\cite{Aubry14}. 
    We have tried to estimate the parameters in a similar manner by fitting a line to the tail of the cdf, however this procedure did not yield satisfactory results.
    In addition, as discussed above, in our case this requires running all $N$ classifiers on all $N$ images, which is prohibitive for large datasets.
    Instead, we have found that a good calibration can be obtained by normalizing the learnt hyperplane $\vec{w}$. In particular, we set 
    %
      \begin{align}
        \label{eq:rescale1}
        \alpha_j &= \dfrac{1}{||\vec{w}_j||},\\ 
        \label{eq:rescale2}
        \beta_j &=  -b_j \alpha_j, 
      \end{align}
    %
    where $\vec{w}_j$ and $b_j$ are the parameters of the learnt SVM hyper-plane for location $j$ and $||\vec{w}||$ is the $L_2$ norm of $\vec{w}$ . Given this choice of $\alpha_j$ and $\beta_j$ the calibrated classification score~\eqref{eq:linear2} reduces to
    %
      \begin{align}
        f_j(\vec{q}) = \dfrac{1}{||\vec{w}_j||}\vec{w}_j^T\vec{q} = \vec{\widetilde{w}}_j^T\vec{q}.
        \label{eq:linear3}
      \end{align}
    %
    The intuition is that when $\vec{q}$ is $L_2$ normalized, equation~\eqref{eq:linear3} is equivalent to computing the normalized dot-product between vectors $\vec{q}$ and $\vec{w}$.       This was found to work well in image retrieval~\cite{Sivic03} or matching whitened HOG descriptors~\cite{Doersch13}.
    In this work we investigate whether this intuition about descriptor matching can be used as a form of calibration for the learnt place-specific classifier.  
    Note  that this form of calibration by normalization is scalable to very large datasets as it (i) requires only $O(N)$ computations offline to pre-compute the calibration parameters for each of the $N$ learnt classifiers (equations~\eqref{eq:rescale1} and~\eqref{eq:rescale2}), and (ii) does not need any additional storage or computation at query time as the calibration parameters can be included in the classifier~\eqref{eq:linear2}. In Appendix we examine the per-exemplar SVM cost and give an additional intuition why calibration by re-normalization works.

@@

  %%%%%%%%%%%%%%%%%%%%%%%%%%
  \subsection{Image datasets}
  %%%%%%%%%%%%%%%%%%%%%%%%%%
    {
      Experiments are performed on two datasets, the Pittsburgh place recognition dataset~\cite{Gronat13} and the Tokyo 24/7 dataset~\cite{Torii2015}. 
    }
    \paragraph{Pittsburgh dataset.}
      {
        The first dataset contains Google Street View panoramas downloaded from the Internet covering an area of $1.3 \times 1.2 \; {\rm km}^2$ of the city of Pittsburgh (U.S.). Similar to~\cite{Chen11}, we generate for each panorama 12 overlapping perspective views corresponding to two different elevation angles $4^\circ$ and $28^\circ$ to capture both the street-level scene and the building fa\c{c}ades. This results in a total of 24 perspective views each with $90^\circ$ FOV and resolution of $960 \times 720$ pixels. In this manner we generate two versions of this dataset. The first version covers a smaller area and contains $25$k perspective images.  The second larger dataset contains $55$k images.
      }

      {
        As a query set with known ground truth GPS positions, we use 8999 panoramas from the Google Street View research dataset~\cite{GoogleResearch}, which cover approximately the same area, but were captured at a different time, and typically depict the same places from different viewpoints and under different illumination conditions.          
        We generate a test query set such that we first select a panorama at random, and second, we generate a perspective image with a random orientation and random elevation pitch. This way we synthesize 4,000 query test images. Both the query and database images are available upon request at~\cite{GronatProjectPage}.
      }

    \paragraph{24/7 Tokyo dataset.}
      {
        The 24/7 Tokyo dataset~\cite{Torii2015} contains Google Street View panoramas downloaded from the Internet covering an area of $1.6 \times 1.6 \; {\rm km}^2$ of the city of Tokyo. The dataset contains 76k perspective views. The query set contains 315 query images from 105 distinct locations captured by different types of camera phones. This dataset is very challenging as each location is captured at three different times: during day, at sunset and during night. 
        The dataset is available upon request at~\cite{Torii15project}.
      }

@@

    %%%%%%%%%%%%%%%%%%%%%%%%%%
    \subsection{Parameters of per-location classifier learning}
    %%%%%%%%%%%%%%%%%%%%%%%%%%
      To learn the exemplar support vector machine for each database image $j$, the positive and negative training data are constructed as follows. 
      The \emph{negative training set} $\mathcal N_j$ is obtained by: (i) finding the set of images with geographical distance greater than $200\;m$; (ii)  sorting the images by decreasing value of similarity to image $j$ measured by the dot product between their respective descriptors; (iii) taking the top $N=500$ ranked images as the negative set. 
      In other words, the negative training data consists of the hard negative images, i.e. those that are similar to image $j$ but are far away from its geographical position, hence, cannot have the same visual content. 
      The \emph{positive training set} $\mathcal P_j$ consist of the descriptor $\vec{x}_j$ of the target image $j$.

      We found that for the bag-of-visual-words representation it was useful to further expand~\cite{Chum07b} positive training set by close by images that view the same scene structures. These images can be identified by geometric verification \cite{Philbin07}  {as follows.}
      {
        We first build a graph where each image in the database represents a node and an edge represents a spatial adjacency in the world. An edge is present if the positions of the two images are within $50m$ of each other.
        Then, we score each edge by the number of geometrically verified matches~\cite{Philbin07}. Finally, we remove edges with score below a threshold of $t_m = 40$ matches. 
        It is worth noting that the graph contains many isolated nodes. This typically indicates that the viewpoint change between two adjacent panoramas is large. For each image in the database, we include between zero and five extra positive examples that are directly connected in the graph.
      }

@@

  %%%%%%%%%%%%%%%%%
  \subsection{Bag-of-visual-words model}
  \label{sec:bow_results}
  %%%%%%%%%%%%%%%%%
%
    Results for the bag-of-visual-words image representation are shown in table~\ref{tab:recallBOW}.  
    Learning per-location classifiers with either calibration method (\emph{p-val} and \emph{w-norm}) clearly improves over the standard bag-of-visual-words baseline (BOW) that does not perform any learning. In addition, both calibration methods significantly improve over the learnt SVM classifiers without any calibration (BOW SVM no calib) underscoring the importance of calibration for the independently learnt per-location classifiers.  
    {
      In table~\ref{tab:recallBOW}, we also compare performance to our implementation of the confuser suppression approach (Conf. supp.) of~\cite{Knopp2010} that, in each database image, detects and removes features that frequently appear at other far-away locations (using parameters $t=3.5$ and $w=70$). The results show an improvement by our method, specially at recall@1.
    }

@@

      We have compared our method to the linear discriminant analysis (LDA)~\cite{Aubry13,Hariharan12,Gharbi12} and whitening~\cite{JegouChum12} baselines. Results are reported on the Pittsburgh $25k$ dataset. The LDA baseline finds a discriminative linear projection of the feature space by minimizing an Euclidean loss rather than hinge loss used in our work. In detail, following~\cite{Aubry13} we have used all available database to learn the covariance matrix and used calibrated LDA score (see~\cite{Aubry13} eq.~11) to obtain a classifier for each database image.  

>>>>

\begin{abstract}
  The aim of this work is to localize a query photograph by finding other images depicting the same place in a large geotagged image database. This is a challenging task due to changes in viewpoint, imaging conditions and the large size of the image database. The contribution of this work is two-fold. First, we cast the place recognition problem as a classification task and use the available geotags to train a classifier for each location in the database in a similar manner to per-exemplar SVMs in object recognition. 
  Second, as only 
  \textcolor{petr}{one or} 
  a few positive training examples are available for each location, we propose two methods to calibrate all the per-location SVM classifiers without the need for additional positive training data.
  The first method relies on p-values from statistical hypothesis testing and uses only the available negative training data. The second method performs an affine calibration by appropriately normalizing the learnt classifier hyperplane and does not need any additional labelled training data. We test the proposed place recognition method with the bag-of-visual-words and Fisher vector image representations suitable for large scale indexing. 
  {  
    Experiments are performed on three datasets: 25,000 and 55,000 geotagged street view images of Pittsburgh, and the 24/7 Tokyo benchmark containing $76,000$ images with varying illumination conditions. The results show improved place recognition accuracy of the learnt image representation over direct matching of raw image descriptors.
  }

@@

  In this work, we also take advantage of geotags as an available form of supervision and investigate whether the place recognition problem can be cast as a classification task.
  {
    Learning visual classifiers has been investigated for landmark recognition~\cite{Li09} where consumer photographs were clustered into landmark classes based on geo-tags. 
    In this work we wish to recognize individual street locations rather than a small number of landmarks, and as a consequence have only a few (1-5) photographs
    %available for each place.   
    \textcolor{petr}{capturing the same location}.
    In particular, we train a classifier {\em for each location on the map} in a similar manner to per-exemplar classification in object recognition~\cite{Malisiewicz11}. 
    \textcolor{petr}{
      By \em{location} we mean a scene captured by a camera, hence using our terminology, a camera heading to the north captures different location than a camera heading to the south, even those cameras share the same GPS location. 
    }
  }

@@

  While learning classifiers for each 
  %place 
  \textcolor{petr}{location} 
  may be appealing, calibrating outputs of the individual classifiers is a critical issue. In object recognition~\cite{Malisiewicz11}, it is addressed in a separate calibration stage on a held-out set of training data.
  This is not possible in the place recognition set-up as only a small number, typically one to five, of positive training images are available for each location (e.g. Street View images viewing the same building facade). To address this issue, we propose two calibration methods. 
  The first method relies on p-values from statistical hypothesis testing and uses only the available negative training data. The second method performs a simple affine calibration by appropriately normalizing the learnt classifiers and does not need any additional labelled calibration examples.

@@

  The task of geo-localizing a given input query photograph has recently received considerable attention. The output can be a coarse geo-localization on the level of continents and cities~\cite{Doersch12,Hays2008,Kalogerakis09} or a name of the depicted landmark~\cite{Li09}.  In this work we focus on visually recognizing the ``same place" by finding an image in geo-tagged database that depicts the same building facade or street-corner as shown in the query~\cite{Chen11,Cummins09,Knopp2010,Schindler07,Torii2013,Zamir10}.

  This {\em visual place recognition} problem is typically treated as large-scale instance-level retrieval~\cite{Cummins09,Chen11,Knopp2010,Schindler07,Torii2013,Zamir10}, where images are represented using local invariant features~\cite{Lowe04} encoded and aggregated into the bag-of-visual-words~\cite{Csurka04,Sivic03} or Fisher vector~\cite{Jegou12} representations. The image database can be further augmented by 3D point clouds~\cite{Klinger13}, automatically reconstructed by large-scale structure from motion (SfM)~\cite{Agarwal-ICCV-2009,Klinger13}, which enables accurate prediction of query image camera position~\cite{Li12,Sattler12}.
  In contrast, in this work we investigate learning a discriminative place-specific image representation. A similar idea has been recently explored in~\cite{Cao13} who learn a graph-based discriminative representation for landmark image collections where typically many images are available for each landmark.
  In this work, we focus on street-level images such as Google Street View, which have greater coverage, but typically only one or a small number of images are  available for each 
  %place
  \textcolor{petr}{location}.  
  To address this issue we learn a discriminative re-weighting of the descriptor specific to each image in the database using per-exemplar support vector machine~\cite{Malisiewicz11}.

  \paragraph{Per-exemplar support vector machines.} 
    The exemplar support vector machines (e-SVM) has been used in a number of visual recognition tasks including category-level recognition~\cite{Malisiewicz11}, cross-domain retrieval~\cite{Shrivastava11}, scene parsing~\cite{Tighe13} or as an initialization  for more complex discriminative clustering models~\cite{Doersch12,Singh12}. The main idea is to train a linear support vector machine (SVM) classifier from a single positive example
    \textcolor{petr}{(or very few positive examples)}
     and a large number of negatives. The intuition is that the resulting weight vector will give a higher weight to the discriminative dimensions of the positive training data point and will down weight dimensions that are non-discriminative with respect to the negative training data.

    The exemplar support vector machine can be learnt at query time where the weight vector is used as a new query image representation~\cite{Shrivastava11}. However, this requires training a new classifier afresh for each query that is computationally very demanding. In this work, similar to~\cite{Malisiewicz11} who learn per-exemplar object category representation, we learn per-exemplar classifiers for each place in the database off-line.  A key advantage is that each per-exemplar classifier is trained independently and hence the learning can be heavily parallelized. The per-exemplar training brings, however, also an important drawback. As each classifier is trained independently a careful calibration of the resulting classifier scores is required~\cite{Malisiewicz11}.

@@

\paragraph{Contributions.} 
  This paper has two main contributions. 
  %
  First, we cast the place recognition problem as a classification task where we use the available geo-tags as a weak form of supervision to train a classifier for each location in the database (section~\ref{sec:classifiers}).
  These classifiers are subsequently used for ranking the database images at query time.

  Second, as only 
  \textcolor{petr}{one or} 
  a few positive training examples are available for each location, we propose two methods to calibrate all the per-location SVM classifiers without the need for additional positive training data. The first method (section~\ref{sec:calibration}) relies on p-values from statistical hypothesis testing. The second method (section~\ref{sec:calibrationRenorm}) performs an affine calibration by appropriately normalizing the learnt decision hyperplane. We also describe a memory efficient classifier representation for the sparse bag-of-visual-word vectors (section~\ref{sec:memory}) and  experimentally demonstrate benefits of the proposed approach (section~\ref{sec:experiments} and~\ref{sec:results}).

%%%%%%%%%%%%%%%%%%%%%%%%
\section{Per-location classifiers for place recognition}
\label{sec:classifiers}
%%%%%%%%%%%%%%%%%%%%%%%%
   We are given an image descriptor $\vec{x}_j$, one for each database image $j$. This representation can be a sparse tf-idf weighted bag-of-visual-words vector~\cite{Sivic03} or a dense compact descriptor such as the Fisher vector (FV)~\cite{Jegou12}. The goal is to learn 
   a score \textcolor{petr}{function}
   $f_j$ for each database image $j$, so that, at test time, given the descriptor $\vec{q}$ of the query image, we can either retrieve the correct target image as the image $j^*$ with the highest score
   % 
   \begin{equation}
   \label{eq:class}
    j^*=\operatorname*{arg\;max}_{j} f_j(\vec{q}) 
   \end{equation}
    %
   \noindent
   or use these scores to rank candidate images and use geometric verification to identify the correct location in an $n$-best list.
   Instead of approaching the problem directly as a large multiclass classification problem, we tackle the problem by learning a per-exemplar linear SVM classifier~\cite{Malisiewicz11}  for each database image $j$.
   Similar to~\cite{Knopp2010}, we use the available geotags to construct the negative set $\mathcal N_j$ for each image $j$. The negative set is constructed so as to concentrate difficult negative examples, i.e.\ from images that are far away from the location of image $j$ and similar to the target image as measured by the dot product between their feature vectors. The details of the construction procedure will be given in section~\ref{sec:experiments}.  The positive set $\mathcal P_j$ is represented by a single positive example
   , which is $\vec{x}_j$ itself. 
   \textcolor{petr}{
    For some locations the positive set can be expanded, details of this procedure will be given in~\ref{sub:parametersOfPerLoc}.
   }

   Each SVM classifier produces a \textcolor{petr}{classification} score $s_j$ which is a priori not comparable with the score of the other classifiers. A calibration of these \textcolor{petr}{classification} scores will therefore be key to convert them to comparable scores $f_j$. This calibration problem is more difficult than usual given that we only have a single positive example and will be addressed in section~\ref{sec:calibration}.

   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
   \subsection{Learning per-location SVM classifiers }
   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
      Each linear SVM classifier %learns 
      {generates} 
      a \textcolor{petr}{classification} score 
      $s_j$ of the form 
      %
      \begin{align}
        s_j(\vec{q})=\vec{q}^T \vec{w}_j+b_j
        \label{eq:linear}
      \end{align}
      %
      \noindent
      where $\vec{w}_j$ is a weight vector re-weighting contributions of individual visual words and $b_j$ is the bias specific for image $j$. Given the training sets $\mathcal P_j$ and $\mathcal N_j$, the aim is to find a vector $\vec{w}_j$ and bias $b_j$ such that the score difference between $\vec{x_j}$ and the closest neighbor from its negative set $\mathcal N_j$ is {\em maximized}.
      Learning the weight vector $\vec{w}_j$ and bias $b_j$ is formulated as a minimization of the convex objective 
      %
      \begin{align}
        \nonumber
        \Omega(\vec{w}_j,b_j)=||\vec{w}_j||^{2}& +C_1\sum_{\vec{x}\in \mathcal P_j}h(\vec{w}_j^T\vec{x}+b_j)   \\
        \label{eq:obj}
                           & +C_2\sum_{\vec{x}\in \mathcal N_j}h(-\vec{w}_j^T\vec{x}-b_j), 
      \end{align}
      %
      \noindent
      where the first term is the regularizer, the second term is the loss on the positive training data weighted by scalar parameter $C_1$, and the third term is the loss on the negative training data weighted by scalar parameter $C_2$.   
      This is a standard SVM formulation \eqref{eq:obj}, also used in exemplar-SVM~\cite{Malisiewicz11}.
      In our case $h$ is the squared hinge loss, which we found to work better in our setting than the standard hinge-loss. Parameters $\vec{w}_j$ and $b_j$ are learned separately for each database image $j$ in turn.

@@

      Two other examples of score calibration functions are shown in figure~\ref{fig:3qVSw} in section~\ref{sec:results}. Note that while figure~\ref{fig:calib} illustrates only few points on the cdf, the two plots in figure~\ref{fig:3qVSw} show a complete cdf that contains on the order of 25k data points. Note also that the two cumulative density functions in figure~\ref{fig:3qVSw} are similar but not identical.

@@

  \subsection{Calibration by normalization}
  \label{sec:calibration_by_renormalization}
    Parameters of the affine calibration function~\eqref{eq:affine} could be learnt from negative training data in a similar manner to, for example,~\cite{Aubry14}. 
    We have tried to estimate the parameters in a similar manner by fitting a line to the tail of the cdf, however this procedure did not yield satisfactory results.
    In addition, as discussed above, in our case this requires running all $N$ classifiers on all $N$ images, which is prohibitive for large datasets.
    Instead, we have found that a good calibration can be obtained by normalizing the learnt hyperplane $\vec{w}$. In particular, we set 
    %
      \begin{align}
        \label{eq:rescale1}
        \alpha_j &= \dfrac{1}{||\vec{w}_j||},\\ 
        \label{eq:rescale2}
        \beta_j &=  -b_j \alpha_j, 
      \end{align}
    %
    where $\vec{w}_j$ and $b_j$ are the parameters of the learnt SVM hyper-plane for location $j$ and $||\vec{w}||$ is the $L_2$ norm of $\vec{w}$.

    Given this choice of $\alpha_j$ and $\beta_j$ 
    \textcolor{petr}{
      the bias term in eq. \eqref{eq:linear2} cancels out and
    }
    the calibrated classification score~\eqref{eq:linear2} reduces to
    %
      \begin{align}
        f_j(\vec{q}) = \dfrac{1}{||\vec{w}_j||}\vec{w}_j^T\vec{q} = \vec{\widetilde{w}}_j^T\vec{q}.
        \label{eq:linear3}
      \end{align}
    %
    % The intuition is that when $\vec{q}$ is $L_2$ normalized, equation~\eqref{eq:linear3} is equivalent to computing the normalized dot-product between vectors $\vec{q}$ and $\vec{w}$. 
    \textcolor{petr}{
      Since $\vec{q}$ is $L_2$ normalized, the reason of particular choice of $\alpha_j$ and $\beta_j$ is that the equation equation~\eqref{eq:linear3} is equivalent to computing the normalized dot-product between vectors $\vec{q}$ and $\vec{w}$.
    }
    This was found to work well in image retrieval~\cite{Sivic03} or matching whitened HOG descriptors~\cite{Doersch13}.
    In this work we investigate whether this intuition about descriptor matching can be used as a form of calibration for the learnt place-specific classifier.

    Note  that this form of calibration by normalization is scalable to very large datasets as it (i) requires only $O(N)$ computations offline to pre-compute the calibration parameters for each of the $N$ learnt classifiers (equations~\eqref{eq:rescale1} and~\eqref{eq:rescale2}), and (ii) does not need any additional storage or computation at query time as the calibration parameters can be included in the classifier~\eqref{eq:linear2}. In Appendix we examine the per-exemplar SVM cost and give an additional intuition why calibration by re-normalization works.

@@

  %%%%%%%%%%%%%%%%%%%%%%%%%%
  \subsection{Image datasets}
  %%%%%%%%%%%%%%%%%%%%%%%%%%
    {
      Experiments are performed on two datasets, the Pittsburgh place recognition dataset~\cite{Gronat13} and the Tokyo 24/7 dataset~\cite{Torii2015}. 
    }
    \paragraph{Pittsburgh dataset.}
      The first dataset contains Google Street View panoramas downloaded from the Internet covering an area of $1.3 \times 1.2 \; {\rm km}^2$ of the city of Pittsburgh (U.S.). Similar to~\cite{Chen11}, we generate for each panorama 12 overlapping perspective views corresponding to two different elevation angles $4^\circ$ and $28^\circ$ to capture both the street-level scene and the building fa\c{c}ades. This results in a total of 24 perspective views each with $90^\circ$ FOV and resolution of $960 \times 720$ pixels. In this manner we generate two versions of this dataset. The first version covers a smaller area and contains $25$k perspective images.  The second larger dataset contains $55$k images.

      As a query set with known ground truth GPS positions, we use 8999 panoramas from the Google Street View research dataset~\cite{GoogleResearch}, which cover approximately the same area, but were captured at a different time, and typically depict the same places from different viewpoints and under different illumination conditions.          
      We generate a test query set such that we first select a panorama at random, and second, we generate a perspective image with a random orientation and random elevation pitch. This way we synthesize 4,000 query test images. Both the query and database images are available upon request at~\cite{GronatProjectPage}.

    \paragraph{24/7 Tokyo dataset.}
      The 24/7 Tokyo dataset~\cite{Torii2015} contains Google Street View panoramas downloaded from the Internet covering an area of $1.6 \times 1.6 \; {\rm km}^2$ of the city of Tokyo. The dataset contains 76k perspective views. The query set contains 315 query images from 105 distinct locations captured by different types of camera phones. This dataset is very challenging as each location is captured at three different times: during day, at sunset and during night. 
      The dataset is available upon request at~\cite{Torii15project}.

@@

    %%%%%%%%%%%%%%%%%%%%%%%%%%
    \subsection{Parameters of per-location classifier learning}
    \label{sub:paramsOfPerLoc}
    %%%%%%%%%%%%%%%%%%%%%%%%%%
      To learn the exemplar support vector machine for each database image $j$, the positive and negative training data are constructed as follows. 
      The \emph{negative training set} $\mathcal N_j$ is obtained by: (i) finding the set of images with geographical distance greater than $200\;m$; (ii)  sorting the images by decreasing value of similarity to image $j$ measured by the dot product between their respective descriptors; (iii) taking the top $N=500$ ranked images as the negative set. 
      In other words, the negative training data consists of the hard negative images, i.e. those that are similar to image $j$ but are far away from its geographical position, hence, cannot have the same visual content. 
      The \emph{positive training set} $\mathcal P_j$ consist of the descriptor $\vec{x}_j$ of the target image $j$.

      We found that for the bag-of-visual-words representation it was useful to further expand~\cite{Chum07b} positive training set by close by images that view the same scene structures. These images can be identified by geometric verification \cite{Philbin07} as follows.
      %
      We first build a graph where each image in the database represents a node and an edge represents a spatial adjacency in the world. An edge is present if the positions of the two images are within $50m$ of each other.
      Then, we score each edge by the number of geometrically verified matches~\cite{Philbin07}. Finally, we remove edges with score below a threshold of $t_m = 40$ matches. 
      It is worth noting that the graph contains many isolated nodes. This typically indicates that the viewpoint change between two adjacent panoramas is large. For each image in the database, we include between zero and five extra positive examples that are directly connected in the graph.

@@

  %%%%%%%%%%%%%%%%%
  \subsection{Bag-of-visual-words model}
  \label{sec:bow_results}
  %%%%%%%%%%%%%%%%%
%
    Results for the bag-of-visual-words image representation are shown in table~\ref{tab:recallBOW}.  
    Learning per-location classifiers with either calibration method (\emph{p-val} and \emph{w-norm}) clearly improves over the standard bag-of-visual-words baseline (BOW) that does not perform any learning. In addition, both calibration methods significantly improve over the learnt SVM classifiers without any calibration (BOW SVM no calib) underscoring the importance of calibration for the independently learnt per-location classifiers.

    In table~\ref{tab:recallBOW}, we also compare performance to our implementation of the confuser suppression approach (Conf. supp.) of~\cite{Knopp2010} that, in each database image, detects and removes features that frequently appear at other far-away locations (using parameters $t=3.5$ and $w=70$).  
    % R1:
    %The results show an improvement by our method, specially at recall@1.
    \textcolor{petr} {Our results show an improvements, specially at recall@1.}

@@

      We have compared our method to the linear discriminant analysis (LDA)~\cite{Aubry13,Hariharan12,Gharbi12} and whitening~\cite{JegouChum12} baselines. Results are reported on the Pittsburgh 25k dataset. The LDA baseline finds a discriminative linear projection of the feature space by minimizing an Euclidean loss, rather than the hinge loss used in our work. In detail, following~\cite{Aubry13} we have used all available data to learn the covariance matrix and used the calibrated LDA score (see~\cite{Aubry13} eq.~11) to obtain a classifier for each database image.  
